### 2017-05-17 paper/content/parrows.tex
initial

\subsubsection{Multicore Haskell} \label{sec:parrows:multicore}
The Multicore Haskell implementation of this class is implemented in a straightforward manner by using listApp from appendix \ref{utilfns} combined with the \code{withStrategy :: Strategy a -> a -> a} and \code{pseq :: a -> b -> b} combinators from Multicore Haskell, where \code{withStrategy} is the same as \code{using :: a -> Strategy a -> a} but with flipped parameters.
\begin{figure}[h]
\begin{lstlisting}[frame=htrbl]
instance (NFData b, ArrowApply arr, ArrowChoice arr) =>
	ArrowParallel arr a b () where
    	parEvalN _ fs =
       		listApp fs >>>
        	arr (withStrategy (parList rdeepseq)) &&& arr id >>>
        	arr (uncurry pseq)
\end{lstlisting}% $ %% formatting
\caption{Fully evaluating ArrowParallel instance for the Multicore Haskell backend}
\label{fig:ArrowParallelMulticoreRdeepseq}
\end{figure}
For most cases a fully evaluating version like in Fig.~\ref{fig:ArrowParallelMulticoreRdeepseq} would probably suffice, but as the Multicore Haskell interface allows the user to specify the level of evaluation to be done via the \code{Strategy} interface, we want to the user not to lose this ability because of using our API. We therefore introduce the \code{Conf a} data-type that simply wraps a \code{Strategy a} (Fig.~\ref{fig:confa}). We can't directly use the \code{Strategy a} type here as GHC (at least in the versions used for development in this paper) does not allow type synonyms in type class instances.
\begin{figure}[h]
\begin{lstlisting}[frame=htrbl]
data Conf a = Conf (Strategy a)
\end{lstlisting}
\caption{Definition of Conf a}
\label{fig:confa}
\end{figure}
To get our configurable \code{ArrowParallel} instance, we simply unwrap the strategy and pass it to \code{parList} as in the fully evaluating version (Fig.~\ref{fig:ArrowParallelMulticoreConfigurable}).
\begin{figure}[h]
\begin{lstlisting}[frame=htrbl]
instance (NFData b, ArrowApply arr, ArrowChoice arr) =>
	ArrowParallel arr a b (Conf b) where
    	parEvalN (Conf strat) fs =
        	listApp fs >>>
        	arr (withStrategy (parList strat)) &&& arr id >>>
        	arr (uncurry pseq)
\end{lstlisting}
\caption{Configurable ArrowParallel instance for the Multicore Haskell backend}
\label{fig:ArrowParallelMulticoreConfigurable}
\end{figure}
\subsubsection{Par Monad}
The ParMonad implementation (Fig.~\ref{fig:ArrowParallelParMonad}) makes use of Haskells laziness and ParMonad's \code{spawnP :: NFData a =>\ \ a -> Par (IVar a)} function. The latter forks away the computation of a value and returns an \code{IVar} containing the result in the \code{Par} monad.

>>>>

\subsubsection{Multicore Haskell} \label{sec:parrows:multicore}
The Multicore Haskell implementation of this class is implemented in a straightforward manner by using listApp from appendix \ref{utilfns} combined with the \code{withStrategy :: Strategy a -> a -> a} and \code{pseq :: a -> b -> b} combinators from Multicore Haskell, where \code{withStrategy} is the same as \code{using :: a -> Strategy a -> a} but with flipped parameters.
\begin{figure}[h]
\begin{lstlisting}[frame=htrbl]
instance (NFData b, ArrowApply arr, ArrowChoice arr) =>
	ArrowParallel arr a b () where
    	parEvalN _ fs =
       		listApp fs >>>
        	arr (withStrategy (parList rdeepseq)) &&& arr id >>>
        	arr (uncurry pseq)
\end{lstlisting}% $ %% formatting
\caption{Fully evaluating ArrowParallel instance for the Multicore Haskell backend}
\label{fig:ArrowParallelMulticoreRdeepseq}
\end{figure}
For most cases a fully evaluating version like in Fig.~\ref{fig:ArrowParallelMulticoreRdeepseq} would probably suffice, but as the Multicore Haskell interface allows the user to specify the level of evaluation to be done via the \code{Strategy} interface, we want to the user not to lose this ability because of using our API. We therefore introduce the \code{Conf a} data-type that simply wraps a \code{Strategy a} (Fig.~\ref{fig:confa}). We can't directly use the \code{Strategy a} type here as GHC (at least in the versions used for development in this paper) does not allow type synonyms in type class instances.
\begin{figure}[h]
\begin{lstlisting}[frame=htrbl]
data Conf a = Conf (Strategy a)
\end{lstlisting}
\caption{Definition of Conf a}
\label{fig:confa}
\end{figure}
To get our configurable \code{ArrowParallel} instance, we simply unwrap the strategy and pass it to \code{parList} like in the fully evaluating version (Fig.~\ref{fig:ArrowParallelMulticoreConfigurable}).
\begin{figure}[h]
\begin{lstlisting}[frame=htrbl]
instance (NFData b, ArrowApply arr, ArrowChoice arr) =>
	ArrowParallel arr a b (Conf b) where
    	parEvalN (Conf strat) fs =
        	listApp fs >>>
        	arr (withStrategy (parList strat)) &&& arr id >>>
        	arr (uncurry pseq)
\end{lstlisting}
\caption{Configurable ArrowParallel instance for the Multicore Haskell backend}
\label{fig:ArrowParallelMulticoreConfigurable}
\end{figure}
\subsubsection{Par Monad}
The ParMonad implementation (Fig.~\ref{fig:ArrowParallelParMonad}) makes use of Haskells laziness and ParMonad's \code{spawnP :: NFData a =>\ \ a -> Par (IVar a)} function. The latter forks away the computation of a value and returns an \code{IVar} containing the result in the \code{Par} monad.

### 2017-05-17 paper/content/syntacticSugar.tex
initial

\subsection{Syntactic Sugar} \label{syntacticSugar}
For basic arrows, we have the \code{(***)} combinator (Fig.~\ref{fig:***Img},~\ref{fig:***}) which allows us to combine two arrows \code{arr a b} and \code{arr c d} into an arrow \code{arr (a, c) (b, d)} which does both computations at once. This can easily be translated into a parallel version \code{(|***|)} (Fig.~\ref{fig:|***|}) with the use of \code{parEval2}, but for this we require a backend which has an implementation that does not require any configuration (hence the \code{()} as the conf parameter in Fig.~\ref{fig:|***|}).
\begin{figure}[h]
\begin{lstlisting}[frame=htrbl]
(|***|) :: (ArrowChoice arr, ArrowParallel arr (Either a c) (Either b d) ())) =>
	arr a b -> arr c d -> arr (a, c) (b, d)
(|***|) = parEval2 ()
\end{lstlisting}
\caption{Definition of (|***|) - the parallel version of (***)}
\label{fig:|***|}
\end{figure}
% With this we can analogously to the serial \code{&&&}
We define the parallel \code{(|\&\&\&|)} (Fig.~\ref{fig:|&&&|}) in a similar manner to its sequential pendant \code{(\&\&\&)} (Fig.~\ref{fig:&&&Img},~\ref{fig:&&&}).
\begin{figure}[h]
\begin{lstlisting}[frame=htrbl]
(|&&&|) :: (ArrowChoice arr, ArrowParallel arr (Either a a) (Either b c) ()) =>
	arr a b -> arr a c -> arr a (b, c)
(|&&&|) f g = (arr $ \a -> (a, a)) >>> f |***| g
\end{lstlisting} % $ %% formatting
\caption{Definition of (|\&\&\&|) - the parallel version of (\&\&\&)}
\label{fig:|&&&|}
\end{figure}

>>>>

\subsubsection{Syntactic Sugar} \label{syntacticSugar}
For basic arrows, we have the \code{(***)} combinator (Fig.~\ref{fig:***Img},~\ref{fig:***}) which allows us to combine two arrows \code{arr a b} and \code{arr c d} into an arrow \code{arr (a, c) (b, d)} which does both computations at once. This can easily be translated into a parallel version \code{(|***|)} (Fig.~\ref{fig:|***|}) with the use of \code{parEval2}, but for this we require a backend which has an implementation that does not require any configuration (hence the \code{()} as the conf parameter in Fig.~\ref{fig:|***|}).
\begin{figure}[h]
\begin{lstlisting}[frame=htrbl]
(|***|) :: (ArrowChoice arr, ArrowParallel arr (Either a c) (Either b d) ())) =>
	arr a b -> arr c d -> arr (a, c) (b, d)
(|***|) = parEval2 ()
\end{lstlisting}
\caption{Definition of (|***|) - the parallel version of (***)}
\label{fig:|***|}
\end{figure}
% With this we can analogously to the serial \code{&&&}
We define the parallel \code{(|\&\&\&|)} (Fig.~\ref{fig:|&&&|}) in a similar manner to its sequential pendant \code{(\&\&\&)} (Fig.~\ref{fig:&&&Img},~\ref{fig:&&&}).
\begin{figure}[h]
\begin{lstlisting}[frame=htrbl]
(|&&&|) :: (ArrowChoice arr, ArrowParallel arr (Either a a) (Either b c) ()) =>
	arr a b -> arr a c -> arr a (b, c)
(|&&&|) f g = (arr $ \a -> (a, a)) >>> f |***| g
\end{lstlisting} % $ %% formatting
\caption{Definition of (|\&\&\&|) - the parallel version of (\&\&\&)}
\label{fig:|&&&|}
\end{figure}

### 2017-05-17 paper/content/topologySkeletons.tex
initial

\FloatBarrier
\section{Topological Skeletons}
\label{sec:topology-skeletons}
Even though many algorithms can be expressed by parallel maps, some problems require more sophisticated skeletons. The Eden library leverages this problem and already comes with more predefined skeletons, among them a \code{pipe}, a \code{ring} and a \code{torus} implementation \cite{Loogen2012, eden_skel_topology}. These seem like reasonable candidates to be ported to our arrow based parallel Haskell to showcase that we can express such skeletons with Parallel Arrows as well.

>>>>

\FloatBarrier
\section{Topological Skeletons}
\label{sec:topology-skeletons}
Even though many algorithms can be expressed by parallel maps, some problems require more sophisticated skeletons. The Eden library leverages this problem and already comes with more predefined skeletons, among them a \code{pipe}, a \code{ring} and a \code{torus} implementation \cite{Loogen2012, eden_skel_topology}. These seem like reasonable candidates to be ported to our arrow based parallel Haskell to showcase that we can express more sophisticated skeletons with Parallel Arrows as well.
